---
title: "Visualization 01 Challenge"
author: "Maya Gann-Bociek & Matthew Schuelke, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Introduction

NHANES dataset is survey data collected by the US National Center for Health Statistics (NCHS) which has conducted a series of health and nutrition surveys since the early 1960's. It was included in NHANES package.

Here we are going to plot Height against Age with Male and Female seperately.

## Tasks

1. Load the dependencies
2. Load the dataset "nhanes_c1.csv" into your environment
3. Explore the Data
4. Plot the Data

## Load Dependencies

Load the following librarys:

* ggplot2
* readr 
* here

```{r package-load-1}

```

## Load Data

Load in data "nhanes_c1.csv"

```{r}

```

## Explore the Data

Print loaded dataset to explore your data

```{r}

```

## Plot the Data

Try to:

* create scatterplots of 'height' against 'ageMonths'
* Split the plot to 2 facet: one female, one male
* add losses line to each facet
* set appropriate x and y labels
* choose a theme for this plot
 
```{r}

```
